\texttt{Taxonomic supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-04-26 doc/incertae_sedis.tex
Update.

The Open Tree of Life project \citep{HinchliffEtAl2015} seeks to build
a comprehensive supertree \citep[see][]{redelings2017supertree}
covering all of life.
The approach is to combine information in
published phylogenies with a comprehensive taxonomy that supplies
taxon names.
The supertree method has been used with the Open Tree of
Taxonomy \citep[OTT hereafter, see][]{rees2017automated}, but 
could be used in more general contexts.
The taxonomy is a crucial input for several
reasons: it covers a wide range of species, its list of names allows
for alignment of different phylogenetic estimates to each other, and
it provides names for clades.
Only a small proportion of known species
have been included in a phylogenetic analysis, thus a taxonomy is
important for achieving comprehensive coverage of known taxa.
Phylogenetic estimates collected from the published literature often
use different names for the same species.
Lists of synonyms in OTT
(and other taxonomies) allow data curators to ``align'' input
phylogenies to the taxonomy \citep[see][ for discussion of the
curation process that the Open Tree of Life project uses to align phylogenetic
estimates to OTT]{McTavishEtAt2015}.
This alignment is important for recognizing when two different
estimates refer to the same taxon.
Biologists are familiar with names
for ``higher'' taxa (taxa above the species rank).
Thus, naming clades
in the supertree makes the tree easier to navigate and use.

Some taxonomies indicate which taxa are thought to be the result of hybridization, 
but the supertree approach
described here does not consider any hybridization.
Thus the taxonomic
hierarchy can be converted to a tree.
While not all taxonomists
believe that named taxa should correspond to clades \citep[see, for
example,][]{HorandlS2010}, the priniciples of phylogenetic
classification are used by such a large majority of practicing
taxonomists, that we have chosen to treat the taxonomic tree that
mirrors the hierarchy of the taxonomy as an estimate of phylogenetic
relationships.
We refer to labels of taxonomy nodes as ``taxon
names'', and assume that taxon names and taxonomy nodes have a one-to-
one correspondence.
Note that the taxonomy may also contain out-degree
1 nodes.
These nodes correspond to monotypic taxa which contain a
single child of lower taxonomic rank.

The supertree method described by \citet{redelings2017supertree} is
able to use a set of phylogenetic estimates and taxonomy to produce a
supertree with clades names according to the taxonomy.
However, the
method was not able to accommodate the fact that taxonomies frequently
contain nodes with uncertain placement.
These nodes are often labelled
``\emph{incertae sedis}'', which means ``uncertain seat'' in Latin.
The common interpretation of such taxa is that they may not be moved
further towards the root, but may be moved into their sibling taxa.
For example, a genus with a sibling that is a family may be annotated
as ``\emph{incertae familia}'', indicating that it is unclear which
family the genus should be placed in.
\emph{Incertae sedis} taxa
frequently occur when a specimen is identified down to a given
taxonomic level, but no further.
Extinct taxa are often \emph{incertae
sedis}.
OTT is constructed from several source taxonomies, and these
sources include various ways of indicating that the position of a
taxon within the hierarchy is uncertain.
OTT uses a series of
``flags'' to annotate these taxa.
The supertree of \citet{redelings2017supertree} simply pruned these taxa from the
taxonomy and input trees.
Thus, the final supertree lacked taxa which
were \emph{incertae sedis} or the descendants of such taxa.

@@

This approach incorrectly treats the intrusion of an \emph{incertae sedis} taxon
into a sibling taxon as a conflicting with monophyly of the sibling taxon. For
example, in Fig. \ref{fig:Handling-incertae-sedis}, the placement of the
\emph{incertae sedis} taxon $C$ within $A$ and $AB$, is considered
to contradict the monophyly of $A$ and $AB$ and thus to conflict with those taxa.
This has three main effects.  These effects are illustrated in the difference
between the synthesis tree in Fig. \ref{fig:Handling-incertae-sedis}c, which
ignores \emph{incertae sedis} annotations, and the synthesis tree in Fig.
\ref{fig:Handling-incertae-sedis}d, which respects them.

First, this approach results in a less-resolved tree.  While input trees are
free to place incertae sedis taxa within a sibling, the naive approach
interprets this placement as conflict with the sibling.  Since the taxonomy
is ranked below all input trees, conflict is always resolved in favor of the
input trees, and so the sibling in the taxonomy is removed from the synthesis
tree.  For example, in the synthesis tree in Fig.
\ref{fig:Handling-incertae-sedis}c, the placement of taxon $C$ in group $AB$ is
interpreted as a rejection of the monophyly of $AB$. The edge leading to $AB$ is
therefore not incorporated into the synthesis tree.  In contrast, the synthesis
tree in Fig. \ref{fig:Handling-incertae-sedis}d that does not reject the
monophyly of $AB$ retains the grouping of $A$ and $B$.

Second, this approach results in the removal of clade names for all clades that
have an \emph{incertae sedis} taxon placed within them.  For example, in Fig.
\ref{fig:Handling-incertae-sedis}c the names for the clades $A$ and $AB$ are lost.

Third, this approach cannot use information from phylogenies to resolve the
location of uncertain taxa.  For example, the synthesis tree in
\ref{fig:Handling-incertae-sedis}c has lost the label for $A$, and so does not
allow us to note that $C$ has been placed within $A$.  In contrast, the
synthesis tree in Fig. \ref{fig:Handling-incertae-sedis}d retains the label for
$A$, which allows us to resolve taxonomic uncertainty by infering that
\emph{incertae sedis} taxon $C$ has been placed within $A$.

@@

@@

We seek a semantics for supertrees with \emph{incertae sedis }taxa
that satisfy the following properties:
\begin{enumerate}
    \item An \emph{incertae sedis} node may intrude into its siblings and their
descendants.
    \item The semantics of an \emph{incertae sedis} node does not depend on assigned ranks, but only on the tree.
    \item Semantics is based on deriving a rooted (partial) split for each branch of the taxonomy.
      % These could be goals, but maybe discuss them as side-effects:
      % Side effect #1. Don't interdigitate
      % Side effect #2. Allow placing A next to B, and then placing B within A. (See Fig. 6)
\end{enumerate}

@@

\paragraph{(Not using rank information) Rich taxon-specific attachment points}
% Question: why not use rank information?  Answer: well, we could, but lets start simple.
As mentioned above, the \emph{incertae sedis} annotation may mean that a taxon is
unplaced, but only a few placements within the taxonomy are possible.
For example, a genus might be characterized as \emph{incertae sedis}
at the level of the subfamily.
Thus, it could be placed inside any of the
hypothesized sub-families, or within its own subfamily, but it would
be unexpected for it to be placed inside one of the other genera.

In principle this information is easy to incorporate.
Ideally, the \emph{incertae sedis}
annotation some taxon $x$ would be accompanied by a specific
statement about which other taxa might include $x$.
If $\mathcal{P}$ is the set of taxa that are potential ancestors for
incertae sedis taxon $x$, then $\mathcal{L}(x)$ will occur in neither
the include nor the exclude set of any taxon in $\mathcal{P}$.
Unfortunately, utilizing taxon-specific interpretations of the
incertae sedis flag relies on detailed readings of character argumentation
(which is lacking in OTT) or rank-based information (which is frequently
untrustworthy in OTT).
Thus, here we pursue an approach of ascribing a
meaning to the \emph{incertae sedis} label which attempts to capture
the core of the idea that it articulates, but which can be applied
automatically across the taxonomy without additional information.

\paragraph{(Split-based semantics) Equivalence of a taxonomy tree and a list of rooted splits}

@@

As is typically done, we define rooted splits for each edge by noting that
removing the edge would produce two trees. The leaf label sets of these trees
the partition the full leaf set $\mathcal{L}$ into two groups. The include set $\mathcal{I}(e)$ which
does not contain the root, and the exclude set $\mathcal{E}(e)$ which does contain the root.
Such a split may be written
\[ \mathcal{I}(e)|\bullet\mathcal{E}(e).\]
where the $\bullet$ indicates the root. The include set is the set of leaves that are
descendants of the daughter node of the edge. The exclude set is the set of
leaves that are not descendants of that node. If no taxa are \emph{incertae sedis},
then the exclude set for a node is just the total tip set minus the include set
for the node: $\mathcal{E}(e)  =\mathcal{L}-\mathcal{I}(e)$. For a node $n$ on the tip-ward side of an edge $e$, we
may also write $\mathcal{I}(n)$ for $\mathcal{I}(e)$, and $\mathcal{E}(n)$ for $\mathcal{E}(e)$. For any two rooted splits $A=A_{1}|\bullet A_{2}$
and $B=B_{1}|\bullet B_{2}$ , we say that $A$ \emph{displays }$B$ if $B_{1}\subseteq A_{1}$ and $B_{2}\subseteq A_{2}$.

@@

\paragraph{(Split-based semantics) Information content of the \emph{incertae sedis} annotation}

In terms of rooted split interpretation, an \emph{incertae sedis} taxon differs
from a normal taxon because its leaves should not automatically be mentioned in
the exclude set of all of the edges that does not descend from. If a taxon
rooted at node $n$ is marked as \emph{incertae sedis}, then the leaf labels that
are descendants of $n$ will still be in the include set for $n$ and all its ancestors,
but these labels will not necessarily be the exclude set for all other
nodes/edges. They may simply be omitted in some rooted split representations,
resulting in ``partial splits'' -- those which do not necessarily mention all of
the taxa. Thus, when interpreted correctly the \emph{incertae sedis} annotations
are reducing the amount of information that the taxonomy brings to the supertree
problem.

@@

In order to apply supertree methods to rooted trees with \emph{incertae sedis}
taxa, we first propose a semantics of \emph{incertae sedis }in terms of rooted
splits.

\subsubsection{semantics \#1a: The original, simplest semantics}
\begin{align}
    \mathcal{E}(n) & = \mathcal{E}(parent(n))\cup\left[\mathcal{I}(s)\big|s\in
           siblings(n), s \text{ not \emph{incertae sedis}}\right].\label{eq:exclude-set-formula-1a}
\end{align}
This semantics \emph{does} allow interdigitation.

\subsubsection{semantics \#1b: modified to disallow inter-digitating
\emph{incertae sedis} siblings}

An \emph{incertae sedis} taxon can be moved into any of the
descendants of its siblings.
We seek to represent this by constructing
modified splits for each branch of the taxonomy tree.
The include sets
of these splits remain unchanged, but we construct reduced exclude
sets in order to allow \emph{incertae sedis} taxa to intrude into
their sibling taxa.

The reduced exclude set $\mathcal{E}(n)$ for node $n$ should therefore
not contain the descendants of its siblings which are marked
\emph{incertae }sedis\emph{,} but should contain the descendants of
its other siblings\emph{.
}We additionally stipulate that it is not
possible for \emph{incertae sedis} taxa to intrude into their
\emph{incertae sedis} siblings.
(This is not the only way to construct
a split-based semantics for \emph{incertae sedis}, but it avoids
certain complications, as noted in the discussion.) We thus define
$\mathcal{I}(m,n)$ to indicate descendants of $m$ that are excluded
from $n$.
Thus:
\begin{align*}
\mathcal{I}(m,n) & =
    \begin{cases}
        \emptyset & \text{if }n\text{ is \emph{incertae sedis}\,and }m\text{
is not}\\
        \mathcal{I}(m) & \text{otherwise}
    \end{cases}
\end{align*}

Additionally, the exclude set $\mathcal{E}(n)$ should contain the children of
all the ancestors of $n$, unless those children are \emph{incertae sedis}.
To avoid traversing all ancestors separately for each node $n$, we note that the
excluded children of any ancestors will be in the exclude set of the parent of
$n$ already. Therefore, we can write the exclude set of a node $n$ in terms of
the exclude set of its parent:
\begin{align}
    \mathcal{E}(n) & = \mathcal{E}(parent(n))\cup\left[\mathcal{I}(m,n)\big|m\in
           siblings(n)\right].\label{eq:exclude-set-formula-1}
\end{align}
This formula allows us to compute exclude sets via a pre-order traversal on the
taxonomy tree with exclude set for the root node defined to be $\emptyset$.

\textbf{Problem:} how about the goal that we should be able to get the
same outcome by moving a taxon up one level and marking it as incertae
sedis? This might require that moving A next to B, and then moving B
into A is allowed.

\textbf{Problem:} should we be able to move incertae sedis taxa down
into incertae sedis taxa? If not, then we must change the recursion.

\subsubsection{An alternative semantics that allows placing an incertae sedis group within an incertae-sedis great-uncle}

1. Inter-digitation

@@

@@

In our semantics of incertae sedis taxa, it is possible for to place
one \emph{incertae sedis} sibling inside another.
A special case of
this is when a higher-rank incertae sedis taxon $A$ is placed as
sister to an incertae sedis taxon $B$, and then $B$ is placed within
$A$ (Figure \#).

@@

>>>>

The Open Tree of Life project \citep{HinchliffEtAl2015} seeks to build a
comprehensive supertree \citep[see][]{redelings2017supertree} covering all of
life. The approach is to combine information in published phylogenies with a
comprehensive taxonomy that supplies taxon names. The supertree method has
been used with the Open Tree Taxonomy \citep[OTT hereafter, see][]{rees2017automated},
but could be used in more general contexts. The taxonomy is a crucial input for
several reasons: it covers a wide range of species, its list of names allows
for alignment of different phylogenetic estimates to each other, and it provides
names for clades. Only a small proportion of known species have been included in
a phylogenetic analysis, thus a taxonomy is important for achieving
comprehensive coverage of known taxa. Phylogenetic estimates collected from the
published literature often use different names for the same species. Lists of
synonyms in OTT (and other taxonomies) allow data curators to ``align'' input
phylogenies to the taxonomy \citep[see][ for discussion of the curation process
that the Open Tree of Life project uses to align phylogenetic estimates to OTT]{McTavishEtAt2015}.
This alignment is important for recognizing when two different estimates refer
to the same taxon. Biologists are familiar with names for ``higher'' taxa (taxa
above the species rank). Thus, naming clades in the supertree makes the tree
easier to navigate and use.

Some taxonomies indicate which taxa are thought to be the result of
hybridization, but the supertree approach described here does not consider any
hybridization. Thus the taxonomic hierarchy can be converted to a tree. While
not all taxonomists believe that named taxa should correspond to clades \citep[see,
for example,][]{HorandlS2010}, the priniciples of phylogenetic classification
are used by such a large majority of practicing taxonomists, that we have chosen
to treat the taxonomic tree that mirrors the hierarchy of the taxonomy as an
estimate of phylogenetic relationships. We refer to labels of taxonomy nodes as
``taxon names'', and assume that taxon names and taxonomy nodes have a
one-to-one correspondence. Note that the taxonomy may also contain out-degree
1 nodes. These nodes correspond to monotypic taxa which contain a single child
of lower taxonomic rank.

The supertree method described by \citet{redelings2017supertree} is able to use
a set of phylogenetic estimates and taxonomy to produce a supertree with clades
named according to the taxonomy. However, the method was not able to accommodate
the fact that taxonomies frequently contain nodes with uncertain placement.
These nodes are often labelled ``\emph{incertae sedis}'', which means
``uncertain seat'' in Latin. The common interpretation of such taxa is that
they may not be moved further towards the root, but may be moved into their
sibling taxa. For example, a genus with a sibling that is a family may be
annotated as ``\emph{incertae familia}'', indicating that it is unclear which
family the genus should be placed in. \emph{Incertae sedis} taxa frequently
occur when a specimen is identified down to a given taxonomic level, but no
further. Extinct taxa are often \emph{incertae sedis}. OTT is constructed from
several source taxonomies, and these sources include various ways of indicating
that the position of a taxon within the hierarchy is uncertain. OTT uses a
series of ``flags'' to annotate these taxa. The supertree of
\citet{redelings2017supertree} simply pruned these taxa from the taxonomy and
input trees. Thus, the final supertree lacked taxa which were \emph{incertae
sedis} or the descendants of such taxa.

@@

This approach incorrectly treats the intrusion of an \emph{incertae sedis} taxon
into a sibling taxon as a conflicting with monophyly of the sibling taxon. For
example, in Fig. \ref{fig:Handling-incertae-sedis}, the placement of the
\emph{incertae sedis} taxon $C$ within $A$ and $AB$ is considered
to contradict the monophyly of $A$ and $AB$ and thus to conflict with those taxa.
This has three main effects.  These effects are illustrated in the difference
between the synthesis tree in Fig. \ref{fig:Handling-incertae-sedis}c, which
ignores \emph{incertae sedis} annotations, and the synthesis tree in Fig.
\ref{fig:Handling-incertae-sedis}d, which respects them.

First, the naive approach results in a less-resolved tree.  While input trees are
free to place incertae sedis taxa within a sibling, the naive approach
interprets this placement as conflict with the sibling.  Since the taxonomy
is ranked below all input trees, conflict is always resolved in favor of the
input trees, and so the sibling in the taxonomy is removed from the synthesis
tree.  For example, in the synthesis tree in Fig.
\ref{fig:Handling-incertae-sedis}c, the placement of taxon $C$ in group $AB$ is
interpreted as a rejection of the monophyly of $AB$. The edge leading to $AB$ is
therefore not incorporated into the synthesis tree.  In contrast, the synthesis
tree in Fig. \ref{fig:Handling-incertae-sedis}d that does not reject the
monophyly of $AB$ retains the grouping of $A$ and $B$.

Second, the naive approach results in the removal of clade names for all clades that
have an \emph{incertae sedis} taxon placed within them.  For example, in Fig.
\ref{fig:Handling-incertae-sedis}c the names for the clades $A$ and $AB$ are lost.

Third, the naive approach cannot use information from phylogenies to resolve the
location of uncertain taxa.  For example, the synthesis tree in
\ref{fig:Handling-incertae-sedis}c has lost the label for $A$, and so does not
allow us to note that $C$ has been placed within $A$.  In contrast, the synthesis
tree in Fig. \ref{fig:Handling-incertae-sedis}d retains the label for $A$, which
allows us to resolve taxonomic uncertainty by inferring that $C$ has been placed
within $A$.

@@

\subsection{Goals for an \emph{incertae sedis} semantics}

@@

We seek a semantics for supertrees with \emph{incertae sedis }taxa
that satisfy the following properties:
\begin{enumerate}
    \item An \emph{incertae sedis} node may intrude into its siblings and their descendants without conflict.
    \item An \emph{incertae sedis} node has an unrestricted range within the descendants of its sibling taxa.
    \item The semantics of an \emph{incertae sedis} node does not depend on side-information such as taxon ranks, but only on the taxonomy tree.
    \item Semantics is based on deriving a rooted (partial) split for each branch of the taxonomy.
      % These could be goals, but maybe discuss them as side-effects:
      % Side effect #1. Don't interdigitate
      % Side effect #2. Allow placing A next to B, and then placing B within A. (See Fig. 6)
\end{enumerate}

@@

\subsubsection{Unrestricted range and ignoring side information}
The choice to allow each \emph{incertae sedis} taxon to attach anywhere within
the descendants of its siblings requires some explanation, for two reasons.
First, more precise information about where a taxon can attach is easy to
incorporate. If $\mathcal{P}$ is the set of taxa that are potential ancestors for \emph{incertae
sedis} taxon $x$, then we may simply remove the descendants of $x$ from the exclude
set for each $p\in\mathcal{P}$.  Such an approach would require that the path from $x$ to $p$
go through only other potential ancestors.

Second, as mentioned above, it may be expected that an \emph{incertae sedis} taxon
attach only a specific points within the taxonomy. For example, a genus
might be characterized as \emph{incertae sedis} at the level of the subfamily.
Thus, it could be placed inside any of the hypothesized sub-families, or within
its own subfamily, but it would be unexpected for it to be placed inside one of
the other genera.

Unfortunately, reliable information about the range of valid attachment points
for each \emph{incertae sedis} taxon is lacking.  Ideally, the \emph{incertae
  sedis} annotation for some taxon $x$ would be accompanied by a specific statement
about which other taxa might include $x$.  However, OTT does not provide this
directly.  Constructing a range for each incertae sedis taxon relies on detailed
readings of character argumentation (which is lacking in OTT) or rank-based
information (which is frequently untrustworthy in OTT). Thus, here we pursue
an approach of ascribing a meaning to the \emph{incertae sedis} label which
attempts to capture the core of the idea that it articulates, but which can be
applied automatically across the taxonomy without additional information.

\subsubsection{Split-based semantics}
\paragraph{Equivalence of a taxonomy tree and a list of rooted splits}

@@

\paragraph{Definition of rooted splits}
As is typically done, we define rooted splits for each edge by noting that
removing the edge would produce two trees. The leaf label sets of these trees
the partition the full leaf set $\mathcal{L}$ into two groups. The include set $\mathcal{I}(e)$ which
does not contain the root, and the exclude set $\mathcal{E}(e)$ which does contain the root.
Such a split may be written
\[ \mathcal{I}(e)|\bullet\mathcal{E}(e).\]
where the $\bullet$ indicates the root. The include set is the set of leaves that are
descendants of the daughter node of the edge. The exclude set is the set of
leaves that are not descendants of that node. If no taxa are \emph{incertae sedis},
then the exclude set for a node is just the total tip set minus the include set
for the node: $\mathcal{E}(e)  =\mathcal{L}-\mathcal{I}(e)$. For a node $n$ on the tip-ward side of an edge $e$, we
may also write $\mathcal{I}(n)$ for $\mathcal{I}(e)$, and $\mathcal{E}(n)$ for $\mathcal{E}(e)$. For any two rooted splits $A=A_{1}|\bullet A_{2}$
and $B=B_{1}|\bullet B_{2}$ , we say that $A$ \emph{displays }$B$ if $B_{1}\subseteq A_{1}$ and $B_{2}\subseteq A_{2}$.

@@

@@

An \emph{incertae sedis} taxon $n$ differs from a normal taxon because it can be
moved into the descendants of any of its siblings.  Thus, annotating $n$ as
\emph{incertae sedis} really modifies the meanings of its siblings so that they
no longer exclude $n$.  In terms of rooted splits, we can represent this lack of
exclusion by omitting taxa from the exclude set for a node.  Such a split does
not mention the complete set of tip labels $\mathcal{L}$ and is therefore called a
`partial split'', as opposed to a ``full split'' that contains the entirely of $\mathcal{L}$.
Thus, when interpreted correctly the \emph{incertae sedis} 
annotations are reducing the amount of information that the taxonomy brings to
the supertree problem.

We seek to define modified splits for each branch of the taxonomy tree. The
include sets of these splits remain unchanged, but the exclude sets may be reduced
\emph{incertae sedis} taxa to intrude into their sibling taxa.  Therefore, a
semantics can be given simply in terms of the exclude set for each taxonomy node.

In order to determine how the exclude sets should be modified to account for
\emph{incertae sedis} taxa, we first attempt to express the unmodified exclude
sets in a helpful way.  We note that a node $n$ ordinarily excludes the
descendants of $\mathcal{I}(s)$ for all of its siblings, as well as the siblings
of all of its ancestors.  This leads to the following equation:
\begin{align*}
\mathcal{E}(n) & =\left[\mathcal{I}(s)\big|a\in n+ancestors(n),s\in siblings(a)\right]\\
 & =\left[\mathcal{I}(s)\big|s\in siblings(n)\right]\cup\left[\mathcal{I}(s)\big|a\in ancestors(n),s\in siblings(a)\right]\\
 & =\left[\mathcal{I}(s)\big|s\in siblings(n)\right]\cup\mathcal{E}(parent(n)) \label{eq:exclude-set-formula-traditional}
\end{align*}

When modifying the exclude set for a node $n$, we seek to find all the \emph{incertae
sedis} taxa that may intrude into it. It turns out that the incertae sedis taxa that
might intrude are just the \emph{incertae sedis} siblings of $n$ and all of its ancestors.
\begin{align*}
\mathcal{E}(n) & =\left[\mathcal{I}(s)\big|a\in n+ancestors(n),s\in siblings(a),s\text{ not \emph{incertae sedis}}\right]\\
 & =\left[\mathcal{I}(s)\big|s\in siblings(n),s\text{ not \emph{incertae sedis}}\right]\cup\\
 & \,\,\,\,\,\,\,\,\,\,\,\,\left[\mathcal{I}(s)\big|a\in ancestors(n),s\in siblings(a),s\text{ not \emph{incertae sedis}}\right]\\
 & =\left[\mathcal{I}(s)\big|s\in siblings(n),s\text{ not \emph{incertae sedis}}\right]\cup\mathcal{E}(parent(n)).\label{eq:exclude-set-formula-1a}
\end{align*}
This formula allows us to compute exclude sets via a pre-order traversal on the
taxonomy tree with exclude set for the root node defined to be $\emptyset$.

@@

\subsection{Cases that this semantics can handle}

\subsubsection{Inter-digitation}
1. Inter-digitation

\bdrcomment{Ben: Add figure}

\subsubsection{Placing one \emph{incertae sedis} group within another}
It would be nice if our semantics of incertae sedis taxa would allow us to place
one \emph{incertae sedis} sibling inside another.
A special case of
this is when a higher-rank incertae sedis taxon $A$ is placed as
sister to an incertae sedis taxon $B$, and then $B$ is placed within
$A$ (Figure \#).

\bdrcomment{Ben: Write discussion of Figure \ref{fig:Placing-one-incertae}}

@@

@@

\appendix
\section{Alternative semantics}
\subsubsection{semantics \#1b: modified to disallow inter-digitating
\emph{incertae sedis} siblings}

We additionally stipulate that it is not possible for \emph{incertae sedis} taxa
to intrude into their \emph{incertae sedis} siblings.  We thus define $\mathcal{I}(m,n)$ to
indicate descendants of $m$ that are excluded from $n$. Thus:
\begin{align*}
\mathcal{I}(m,n) & =
    \begin{cases}
        \emptyset & \text{if }n\text{ is \emph{incertae sedis}\,and }m\text{
is not}\\
        \mathcal{I}(m) & \text{otherwise}
    \end{cases}
\end{align*}

We can write the exclude set of a node $n$ in terms of
the exclude set of its parent:
\begin{align}
    \mathcal{E}(n) & = \mathcal{E}(parent(n))\cup\left[\mathcal{I}(m,n)\big|m\in
           siblings(n)\right].\label{eq:exclude-set-formula-1}
\end{align}
\subsubsection{An alternative semantics that allows placing an incertae sedis group within an incertae-sedis great-uncle}
